Kaggle Titanic Training Dataset Description and Munging 
========================================================
### Intial dataset called "data"
```{r setwd, eval=TRUE, echo=FALSE}
setwd('/Users/Teleak/Desktop/Kaggle/RawData') ### set working directory and read in training file
data<-read.csv("train.csv", colClasses=c("factor", "factor", "character", "factor", "numeric", "factor", "factor", "character", "numeric", "character", "character"), strip.white=TRUE, sep=",")
```
******
```{r, libraries, echo=FALSE, results='hide', message=FALSE}
require(plyr)
```
## Description of whole training dataset "data"
*****
#### Dimensions of dataset in rows and columns
```{r, datadim, echo=FALSE, results='markup'}
dim(data)
```
#### Names and classes of data
```{r, nameclass, echo=FALSE, results='markup'}
as.list(sapply(data, class))
```
#### Number of NA's in each column
```{r, nnas, echo=FALSE, results='markup'}
apply(data, 2, function(col) length(which(is.na(col)))) ### shows count of NA values by column
```
#### Number of blank's in each column
```{r, nblanks, echo=FALSE, results='markup'}
apply(data, 2, function(col) length(which(col == ""))) ### shows count of blank values by column
```
#### Summary of whole training set 
```{r, summary, echo=FALSE, results='markup'}
summary(data)
```
### Summary of full data set passenger types by class, sex, and survival
* Full dataset with all NA's and Blanks (none in columns used for calculations).
```{r, datasex, echo=FALSE, results='hide'}
(data.passengers<-ddply(data, .(pclass, sex), summarise, 
			 number = length(sex)
 			))
(data.survived.sum<-ddply(subset(data, data$survived == 1), .(pclass, sex), summarise, 
			 survived = length(sex)
 			))
(data.perished.sum<-ddply(subset(data, data$survived == 0), .(pclass, sex), summarise, 
			 perished = length(sex)
 			))
```
```{r, data.sum.combined, echo=FALSE}
merged.s <- merge(data.passengers, data.survived.sum, by=c('pclass', 'sex' ))
merged <- merge(merged.s, data.perished.sum, by=c('pclass', 'sex'))
ddply(merged, .(pclass, sex, number, survived, perished), summarise, 
			 pclass = pclass,
			 sex = sex, 
			 number = number, 
			 survived = survived, 
			 perc.survive = round(survived/number * 100, digits=2),
	 		 perished = perished, 
	 		 perc.perish = round(perished/number * 100, digits=2)
	 )

```

### Summary of data containing just NA values (those that will be removed)
>__!!! Idea !!!__, may be able to estimate ages from internet
```{r explore.na, echo=FALSE, results='hide'}
has.na<-data[!(complete.cases(data)),] ### subsets data to rows with NA values all of which are in age column (previously determined)
nrow(has.na) ### 177 rows with NA values in age
(na.passengers<-ddply(has.na, .(pclass, sex), summarise, 
			 total = length(sex)
 			))
na.surv <- ddply(subset(has.na, survived == 1), .(pclass, sex), summarise, 
			 survived = length(sex)
 			)
na.perish <- ddply(subset(has.na, survived == 0), .(pclass, sex), summarise, 
			 perished = length(sex)
 			)
merged.na.s <- merge(na.passengers, na.surv, by=c('pclass', 'sex' ))
merge.na <- merge(merged.na.s, na.perish, by=c('pclass', 'sex'))
```
```{r, na.summary, echo=FALSE}
(all.na.sum <- ddply(merge.na, .(pclass, sex, total, survived, perished), summarise, #### combined summary here
			 pclass = pclass,
			 sex = sex, 
			 total = total, 
			 survived = survived, 
			 perc.survive = round(survived/total * 100, digits=2),
	 		 perished = perished, 
	 		 perc.perished = round(perished/total * 100, digits=2)
	 ))
```
### Summary of data after removal of NA's 
****
```{r remove.na, echo=FALSE, results='hide'}
no.na<-data[complete.cases(data),] ### subsets data to rows with NA values all of which are in age column (previously determined)
nrow(no.na) ### 177 rows with NA values in age
(no.na.passengers<-ddply(no.na, .(pclass, sex), summarise, 
			 total = length(sex)
 			))
no.na.surv <- ddply(subset(no.na, survived == 1), .(pclass, sex), summarise, 
			 survived = length(sex)
 			)
no.na.perish <- ddply(subset(no.na, survived == 0), .(pclass, sex), summarise, 
			 perished = length(sex)
 			)
merged.no.na.s <- merge(no.na.passengers, no.na.surv, by=c('pclass', 'sex' ))
merge.no.na <- merge(merged.no.na.s, no.na.perish, by=c('pclass', 'sex'))
```
```{r, no.na.summary, echo=FALSE}
(no.na.sum <- ddply(merge.no.na, .(pclass, sex, total, survived, perished), summarise, #### combined summary here
			 pclass = pclass,
			 sex = sex, 
			 total = total, 
			 survived = survived, 
			 perc.survive = round(survived/total * 100, digits=2),
	 		 perished = perished, 
	 		 perc.perished = round(perished/total * 100, digits=2)
	 ))
```


