### Kaggle Titanic Question

#### Titanic Overview: 
* Sank April 15, 1912
* 1,502 of 2,224 passengers and crew perished

#### Evaluation: 
* Data set split into training and test sets
* For training set, outcome provided ('ground truth')
* Use training set model to generate predictions for test set 

#### Goals
* For each passenger in the test set, one must predict whether or not they survived the sinking
	* 0 == deceased, 1 == survived in data set
	
#### Available Elements (11)
1. survived (1,0)
2. pclass (1, 2, 3)
3. name
4. sex (male or female)
5. age
6. sibsp (number of siblings or spouses aboard -- 0, 1:5, 8)
7. Parch (number of parents/children aboard -- 0:6)
8. fare (passenger fare)
9. cabin 
10. embarked (Port of embarkation)
	* C = Cherbourg; Q = Queenstown; S = Southampton 
	
#### Hypothesis: intuitive assumptions
* Young females with family members aboard (esp. parents/siblings) of higher social status (better accommodation class) were more likely to get a limited lifeboat slot. 
* Survival may have depended upon cabin

#### Hypothesis
* H0 all passengers had an equal chance of survival
* H1 younger female 1st class passengers with family aboard had a higher survival rate. 

#### Exploratory Analysis 
1. % male and female survivors adjusted for % of each in total
2. Relationship to: 
	1. Class of berth
	2. Number of siblings
	3. Survivors Age
	4. Fare paid
	5. Cabin location (look for obvious patterns and then research ship berthing if needed)
	6. Relation to port of embarkation
	
#### Likely Analytical Paths
1. Random Forest Analysis
2. Principal Component Analysis
3. Bayesian Multivariate Prediction 

### Data Description and Munging
********
<!--- Start Analysis JW --->
<!--- set working directory and read in dataset -->
```{r setwd, eval=TRUE, echo=FALSE}
setwd('/Users/Teleak/Desktop/Kaggle/RawData') ### set working directory and read in training file
train<-read.csv("train.csv", colClasses=c("factor", "factor", "character", "factor", "numeric", "factor", "factor", "character", "numeric", "character", "character"), strip.white=TRUE, sep=",")
```
#### Initial data structure
```{r, structure, echo=FALSE, results='markup'}
str(train)
head(train, n=10)
```
#### Columns with NA values
<!-- Figure out NA values -->
```{r, column.NA.count, echo=FALSE, results='markup'}
apply(train, 2, function(col) length(which(is.na(col)))) ### shows count of NA values by column
```
```{r, rows.with.na, echo=FALSE}
has.na<-subset(train, !complete.cases(train)) ### shows rows with NA
na.female<-nrow(subset(has.na, has.na$sex== "female"))
na.male<-nrow(subset(has.na, has.na$sex== "male"))
```

#### Columns with no entry
```{r, noentry}
length(which(train$embarked==""))
length(which(train$cabin==""))
train[train==""] = "unk" ## convert blanks to "unk"
head(train, n=10)

```

#### Summary of missing entities in full training dataset
* `r nrow(has.na) ` missing entities in data frame
* All missing entities in age column
* $\approx$ `r round(na.female/nrow(has.na) * 100)`% of NA values are females
* $\approx$ `r round(na.male/nrow(has.na) * 100)`% of NA values are male

#### Breakdown of males to females prior to sinking
```{r, sex.ratio, echo=FALSE}
n.train<-nrow(train)
males<-nrow(subset(train, train$sex == 'male'))
females<-nrow(subset(train, train$sex == 'female'))
```
* $\approx$ `r round((females/n.train)*100)`% of crew and passengers were female prior to sinking
* $\approx$ `r round((males/n.train)*100)`% of crew and passengers were male prior to sinking


#### Breakdown of males to females after sinking
```{r, survivors, echo=FALSE}
survived<-nrow(subset(train, train[ ,1] == 1))
survived.male <- nrow(subset(train, train[ ,1] == 1 & train$sex == 'male'))
survived.female<-nrow(subset(train, train[ ,1] == 1 & train$sex == 'female'))
```
* $\approx$ `r round((survived.female/survived)*100)`% of the survivors were female
* $\approx$ `r round((survived.male/survived)*100)`% of the survivors were male
